Uncertainty on the galaxy-halo connection for Lyman-$\alpha$ emitters at $z=3.1$

### 2016-03-30 paper/ms.tex
finished revision section 1

We first assume that a dark matter halo can host one detectable LAE at
most.   
This assumption is consistent with theoretical analysis of the
correlation function \citep{Jose2013b} and observations that confirm a
lack of class pairs in LAEs \cite{Bond2009}.  
Then we say that a halo will host a LAE with probability $\focc$
if and only if the halo mass is in the range $\mmin < M_{\mathrm
  h} < \mmax$. 
In what follows we refer to a given set of values $\focc$, $\mmin$,
$\mmax$ as a model, $\mathcal{M}$. We also use a variable \dm to represent
the mass range width, $\dm\equiv\ \log \mmax - \log \mmax$.

The parameter $\focc$ can be thought as the LAE occupation fraction of
dark matter halos.
An astrophysical interpretation of $f_{\rm occ}$ convolves at least
three phenomena: the actual presence of a star forming galaxy in a
halo, the escape fraction of \ly\ radiation and its detectability as
a LAE.   
We do not try to disentangle these complex phenomena.
We instead opt for an arithmentic interpretation by setting $\focc$ as
the ratio between the observed number of LAEs and the number of halos
within the considered mass range, that is $\focc \equiv N_{\rm
  LAEs}/N_{\mathrm{halos}}$.

@@

We do not explore any physical model to disentangle these two effects. This means that our model does not assign a luminosity or escape fraction to each
LAE.
We avoid this in order to maintain theoretical uncertainties to a minimum. 
This flexibility allows us to explore a wide range of possible masses for
the host halos without any strong theoretical prejudice regarding the
details of star formation and \ly\  escape fraction.

We also artificially contaminate our mock catalogs with 22$\%$ randomly distributed data points to mimic the fraction of interlopers in observations \citet{Bielby16} .
On top of that we apply rejection sampling  to our LAE selection taking the transmission function of the NB479 filter used in their observations as a radial selection factor.

In what follows we note by the letter ${\mathcal M}$ a model
defined by a particular choice of the two parameters $M_{\rm min}$, 
$M_{\rm  max}$. For each model  ${\mathcal M}$ we define $\tilde{f}_{\rm occ}$ as the median occupation fraction within the the mock fields and  $\Delta M \equiv \mmax-\mmin$.

\subsection{Exploring and selecting consistent models}
\label{subsec:explore}
We use the angular correlation function to compare the observations with our mock catalogs. 
This is done trough a thorough exploration of the parameter space of the models ${\mathcal M}$ by  means of a Monte Carlo Markov Chain minimization using  the EMCEE python package \citep{emcee2013}.
We put a flat prior on $\log M_{\rm min}$ and $\log \mmax$ to vary between $9.2$ up to $13.4$, corresponding to the halo mass range of the simulation at $z=3$. 
Given that the typical scatter  in $N_{\mathrm halos}$ is about 0.3 dex (i.e. about  a factor of 2), our parameter space is restricted to models where the median number of dark matter halos is larger than  $N_{\rm LAE}/3$. Because of this posibility of having $N_{\mathrm halos}$  smaller than $N_{\rm LAE}$ there will be some models where    $\focc\equiv N_{\rm LAE}/N_{\rm halo}>1$. Those models only reflect the uncertainty in the number density of LAEs due to cosmic variance.

The MCMC exploration is done using a total of 24 seeds and 400
iterations (9600 models) to sample the posterior PDF,  $P(\mmin,
\mmax, f_{\rm occ}| \mathcal{M}) \propto \exp(-\chi_{\mathcal
  M}^2/2)$, with

\begin{equation}
\chi_{\mathcal M}^2=\sum_{\theta}\left[\frac{\left( \rm{ACF}_{\mathcal M}\left(\theta\right) - \rm{ACF}_{\rm obs}\left(\theta\right)\right)^2}{ \sigma_{\rm \mathcal M}^{2}\left(\theta\right) + \sigma_{\rm obs}^{2}\left(\theta\right)}\right]
\end{equation}
%
where  $\rm{ACF}_{\mathcal M}$ and  $\rm{ACF}_{\rm obs}$ are the ACF
of the explored model ${\mathcal M}$ and the observational ACF
reported by \citet{Bielby16} respectively. $\sigma_{\rm \mathcal M}$
is the associated 1-$\sigma$ scatter  of the $\rm{ACF}_{\mathcal M}$
as a product of cosmic variance and  $\sigma_{\rm obs}$ is the
observational error associated to $\rm{ACF}_{\rm obs}$.  The
$\rm{ACF}_{\mathcal M}$s are computed using the Landy \&  Szalay
estimator  \citep{Landy1993}.

>>>>

We first assume that a dark matter halo can host one detectable LAE at
most.   
This assumption is consistent with theoretical analysis of the
correlation function \citep{Jose2013b} and observations that confirm a
lack of class pairs in LAEs \cite{Bond2009}.  
Then we say that a halo will host a LAE with probability $\focc$
if and only if the halo mass is in the range $\mmin < M_{\mathrm
  h} < \mmax$. 
We also use the variable \dm to represent the mass range width,
$\dm\equiv\ \log \mmax - \log \mmax$.

The parameter $\focc$ can be thought as the LAE occupation fraction of
dark matter halos.
An astrophysical interpretation of $f_{\rm occ}$ convolves at least
three phenomena: the actual presence of a star forming galaxy in a
halo, the escape fraction of \ly\ radiation and its detectability as
a LAE.   
We do not try to disentangle these complex phenomena.
We instead opt for an arithmetic interpretation by setting $\focc$ as
the ratio between the observed number of LAEs and the number of halos
within the considered mass range, that is $\focc \equiv N_{\rm
  LAEs}/N_{\mathrm{halos}}$.

For each mock catalog we also randomly remove $22\%$ LAEs to mimic the
effect of contamination in observations \citet{Bielby16}.
On top of that we apply rejection sampling  to our LAE selection
taking the transmission function of the NB479 filter used in their
observations as a radial selection factor.

@@

In what follows we note by the letter ${\mathcal M}$ a model
defined by a particular choice of the two parameters $M_{\rm min}$, 
$M_{\rm  max}$. For each model  ${\mathcal M}$ we define
$\tilde{f}_{\rm occ}$ as the median occupation fraction within the the
mock fields.

\subsection{Exploring and selecting consistent models}
\label{subsec:explore}

We make a thorough parameter space exploration of the models
${\mathcal M}$ by  means of a Markov Chain Monte Carlo technique using
the EMCEE python package \citep{emcee2013}.

We put a flat prior on $\log M_{\rm min}$ and $\log \mmax$ between
$9.2$ up to $13.4$, corresponding to the halo mass range of 
the simulation at $z=3$.  
We restrict the selection to models that give a minimal number denity $N_{\rm{halos}}>N_{\rm
  LAE}/3$.  
This means that it is possible to have $N_{\rm
  {halos}} < N_{\rm LAE}$ and hence $\focc\equiv N_{\rm LAE}/N_{\rm
  halo}>1$.  
This only reflects the uncertainty in the number 
density of LAEs due to cosmic variance.

The MCMC exploration is done using a total of 24 seeds and 400
iterations (9600 models) to sample the posterior PDF,
$P(\mathcal{M}|\mathrm{observations}) \propto \exp(-\chi_{\mathcal
  M}^2/2)$, with the $\chi_{\mathcal{M}}^2$ based on the Angular
Correlation Function (ACF),

\begin{equation}
\chi_{\mathcal M}^2=
\sum_{\theta}\left[\frac{\left( \rm{ACF}_{\mathcal
      M}\left(\theta\right) - \rm{ACF}_{\rm
      obs}\left(\theta\right)\right)^2}{ \sigma_{\rm \mathcal
      M}^{2}\left(\theta\right) + \sigma_{\rm
      obs}^{2}\left(\theta\right)}\right] 
\end{equation}
%
where  $\rm{ACF}_{\mathcal M}$ and  $\rm{ACF}_{\rm obs}$ are the ACF
of the explored model ${\mathcal M}$ and the observational ACF
reported by \citet{Bielby16} respectively. 
$\sigma_{\rm \mathcal M}$ is the associated 1-$\sigma$ scatter  of the
$\rm{ACF}_{\mathcal M}$ as a product of cosmic variance and
$\sigma_{\rm obs}$ is the observational error associated to
$\rm{ACF}_{\rm obs}$.   
We compute the $\rm{ACF}_{\mathcal M}$ using the Landy \&  Szalay
estimator  \citep{Landy1993}.
